---
title: "Bonus Tidyverse CLJ"
output: html_notebook
---


1. Read in `districts.csv` into a tibble named `districts`.

```{r}
library(tidyverse)

districts <- read_csv('../data/districts.csv')
```

2. Notice that the first row corresponds to the whole State of Tennessee. Remove this row and save the result back to `districts`.
```{r}
districts <- districts |> 
  slice(-1)
```

3. How many districts have a proficiency rate of at least 80% for both alg_1 and eng_1?
```{r}
districts |> 
  filter(alg_1 >= 80 & eng_1 >= 80) |> 
  count()
```

4. How many districts have a proficiency rate less than 50% for either alg_1 or eng_1?
```{r}
districts |> 
  filter(alg_1 < 50 | eng_1 < 50) |> 
  count()
```

5. Which district has the lowest graduation rate?
```{r}
districts |> 
  arrange(grad) |> 
  head(1)
```

6. Within the Mid Cumberland region, which district has the highest ACT composite?
```{r}
districts |> 
  filter(region == 'Mid Cumberland') |> 
  arrange(desc(act_composite)) |> 
  head(1)
```

7. Create a scatter plot to compare alg_1 proficiency rates to alg_2 rates. What do you notice? Facet this plot by region. Does anything stand out when you facet the plots?

There seems to be a positive, linear relationship.
```{r}
ggplot(districts, aes(x=alg_1, y=alg_2)) +
  geom_point() +
  facet_wrap(~region)
```

8. When creating this bar chart you may have noticed that some districts have missing enrollment values. For how many districts is this the case?
```{r}
  districts |> 
  filter(is.na(enrollment)) |> 
  nrow()
```

9. A lot of rows are missing additional values. Which district has the largest number of missing values (across all variables)? Hint: you might want to look at rowwise and c_across to answer this question.
```{r}
districts |> 
  rowwise() |> 
  mutate(na_count = sum(is.na(c_across(where(is.numeric))))) |> 
  arrange(desc(na_count)) |> 
  head(2)
```

10. What is the mean graduation rate across all districts? What might be wrong with using just the regular mean to assess average graduation rates?
```{r}
districts |> 
  summarize(mean(grad, na.rm=TRUE))
```

11. Redo the previous question but use a weighted average (`weighted.mean`) graduation across all districts, weighing by enrollment. How much does this change your answer? Can you explain using the data the reason for the big change from using the mean?
```{r}
districts |> 
 drop_na(c(grad, enrollment)) |> 
  summarize(weighted.mean(grad, enrollment))
```

12. Find the unweighted and weighted average graduation rate by district.
```{r}

```

**Continued Exploration and Practice**


13. Read in the school-level testing data for 2014, available [here](https://www.tn.gov/content/dam/tn/education/data/data_2014_state_base.xlsx). You might find the readxl library useful for this task. If you use this library, be sure to look at the `na`argument for the `read_excel` function.
```{r}
library(readxl)

school_test_df <- read_excel('../data/data_2014_state_base.xlsx', na='')

print(school_test_df)
```

To answer the following questions, use "All Students" for the subgroup. 
```{r}
all_students <- school_test_df |> 
  filter(subgroup == 'All Students')
```

14. How many schools have at least 20 percent of students below bsc for Algebra I? Which districts do these schools belong to?
```{r}

```

15. How many schools have at least 20 percent of students below bsc for _both_ Algebra I and English I?
```{r}
all_students |> 
  select()
```

16. Which grade has the highest pct_adv for Algebra I? Plot the average pct_adv per grade level as a bar chart. Make sure that the bars are ordered by grade level.
```{r}

```

17. Find the correlation between pct_adv for Algebra I and pct_adv for Algebra II by school. Create a scatterplot showing Algebra II scores vs. Algebra I scores by school.
```{r}

```

18. Find all schools in Rutherford County that have "High School" in their name. For these schools, create a chart (your choice) showing the differences in pct_below_bsc, pct_bsc, pct_prof, and pct_adv for Algebra I when looking across all subgroups and grades.
```{r}

```

19. Create a function which allows you to select a system_name and which creates a plot to show the differences in pct_below_bsc, pct_bsc, pct_prof, and pct_adv for Algebra I when looking across all subgroups and grades for all schools with "High School" in their names within that system.
```{r}

```

